---
title: "Feature engineering name with 4 approaches to regular expressions"
author: "Shannon Pileggi"
date: 2018-11-12
categories: ["R"]
tags: ["Titanic", "kaggle", "feature engineering", "title", "Name", "regular epressions", "character"]
draft: true
---

```{r, include = F, message = F}
library(tidyverse)
train <- read_csv("C:/Users/spileggi/Google Drive/kaggle/titanic/train.csv")
```

# Overview

In a previous post, we got an overview of Titanic's character variables.  Now let's see what we can do with the `Name` variable; recall, this variable has all unique values!  To get started, let's visually inspect a few values.

```{r}
train %>% 
  select(Name) %>% 
  head(10)
```

In this brief print out, each passenger's title is consistently located between `,` and `.`.  Luckily, this holds true for all observations!  With this consistency, we can somewhat easily extract a title from `Name`. Depending your field, this operation may be referred to something along the lines of _data preparation_ or _feature engineering_. 

Parsing character strings requires knowledge of [regular expressions](https://www.rstudio.com/wp-content/uploads/2016/09/RegExCheatsheet.pdf), which you can think of as a key to unlock string patterns.  Here, the string pattern I want to unlock is 

> ` XXX, YYY. ZZZ`

but without consistent positioning (i.e., the comma isn't always in the 4th position).  Combining the correct key with the correct function will yield the `YYY` content.  Let's demo four ways to tackle the challenge utilizing functions from the [stringr](https://cran.r-project.org/web/packages/stringr/vignettes/stringr.html) package; each methods specifies a different string pattern to match.

```{r}
library(stringr)
```

# Extracting title from name

### First approach

ICYMI, the double bracket in `train[["Name"]]` is used to [extract](https://www.r-bloggers.com/r-accessors-explained/) a named variable vector from a data frame, which has some benefits over the more commonly used dollar sign (i.e., `train$Name`).  Now onward.

The `str_locate` function produces the starting and ending position of a specified pattern.  If we consider the comma to be a pattern, we can figure out where it is located in each name.  Here, the starting and ending value is the same because the comma is only one character.

```{r}
train[["Name"]] %>% 
  str_locate(",") %>%
  head()
```

Knowing this, we can identify the positions of the comma and the period and then extract the text in between.  Some notes here:

* Because `str_locate` function returns a matrix, we use `.` as a [placeholder](https://rafalab.github.io/dsbook/summarizing-data-with-dplyr.html#the-dot-operator) in `.[,1]` to access the first column of values.

* Because `.` is a special character in regular expressions, we use the double backslash in `"\\."` to [escape](https://stringr.tidyverse.org/articles/regular-expressions.html#escaping) it.

```{r}
comma_pos <- 
  train[["Name"]] %>% 
  str_locate(",") %>% 
  .[,1]

period_pos <-
  train[["Name"]] %>% 
  str_locate("\\.") %>% 
  .[,1]
```


Now we can use `str_sub` to extract substrings from the character vector based on their physical position.  To exclude the punctuation and white space, we can add two to the comma position and subtract one from the period position to get the _title_ only.

```{r}
train[["Name"]] %>% 
  str_sub(comma_pos + 2, period_pos - 1) %>% 
  head()
```

Super!

### Second approach

The `str_match` function creates a character matrix for each group matched in the specified pattern.  If we return to our original pattern framework, note that it consists of 5 separate elements:

> `XXX, YYY. ZZZ`

(1) `XXX`, (2) `,`, (3) `YYY`, (4) `.`, and (5) `ZZZ`. With the correct regular expression, `str_match` returns the complete match in addition to all 5 elements (so a matrix with 6 columns).

```{r}
# ---------------------------1---2---3----4---5----                      
str_match("XXX, YYY. ZZZ", "(.*)(, )(.*)(\\.)(.*)")
```

Let's break down this regular expression pattern.  

* the parentheses `()` indicate a grouping
* the 2nd grouping `(, )` only has one modification from the pattern used in the first example - here we include a space after the comma. 
* the 4th `(\\.)` grouping remains unchanged from the first example.
* in the 1st, 3rd, and 5th grouping `(.*)`
  + the period `.` means _any character_
  + the asterisk `*` means _matches at least 0 times_


To execute this, we'll grab the 4th column to catch our title.

```{r}
train[["Name"]] %>% 
  str_match("(.*)(, )(.*)(\\.)(.*)") %>%
  .[,4] %>% 
  head()
```

All right, we got it again!

### Third approach

Lastly, let's use the `str_extract` function to extract matching patterns.  This seems like what we wanted to do all along!   

```{r}
train[["Name"]] %>% 
  str_extract("([A-z]+)\\.") %>%
  head()
```
Let's break down this pattern:

* Here, the bracket `[]`specifies a list of permitted characters.

* Inside the bracket, we specify strings that consist of both upper and lower case letters with `A-z`.

* The `+` outside the bracket signifies "match at least one time", i.e., grab all the letters.

* Finally, ending in `\\.` indicates that our pattern should end in period.

* Putting it all together, this pattern translates to "grab all upper case and lower case letters immediately preceeding a period." 

This pattern is a bit more sophisticated to compose than the previous ones, but it gets right to the point!  This last effort does end in a period, whereas the others do not.  If we wanted to remove the period for consistency, we could use `str_sub` with the `end` argument to specify the position of the last character.


```{r}
train[["Name"]] %>% 
  str_extract("([A-z]+)\\.") %>%
  str_sub(end = -2) %>%
  head()
```


### Fourth approach

As a last effort, we can use `str_replace_all` to replace all matched patterns with null character values.  Here, we specify the pattern and then the replacement string.  

```{r}
train[["Name"]] %>% 
  str_replace_all("(.*, )|(\\..*)", "") %>%
  head()
```
The new regular expression syntax included here is the vertical pipe `|`, which means "or".  In this case, we are specifying that all characters preceeding a comma space (` ,`) and all characters after a period should be replaced with null character values.  

# Re-classifying entries

Now that we figured out how to extract the title, I'll utilize the last method and assign `title` as a variable to the `train` data set using the `mutate` function.

```{r}
train <- 
  train %>% 
  mutate(title = str_replace_all(train[["Name"]],
                                 "(.*, )|(\\..*)", 
                                 ''))
```

Now let's use `count` to get a frequency table of the titles, with the `sort = TRUE` option to arrange the results in descending order.
```{r}
train %>%
  count(title, sort = TRUE)
```

We can see that there are several infrequent titles occuring only one or two times, and so we should re-classify them.  If I was really trying to win, I would try to figure out the historical context and meaning of those titles to create a better classification for them. For now, let's take the easy way out by just re-classifying them to an `other` group.

Fortunately, the `forcats` package has an awesome function that let's us do this quickly: `fct_lump`.  Thus function `lumps` the least frequent values together in an `other` group, and the `n = 6` option specifies to keep the 6 most common values (so the 7th value is `other`).

```{r}
train %>%
  mutate(title2 = fct_lump(title, n = 6)) %>%
  count(title2, sort = TRUE)
```

If you wanted to explicitly re-code the infrequent titles to something more meaningful than `other`, look into [fct_recode](https://r4ds.had.co.nz/factors.html).

