---
title: "An interview presentation inspired by the R community"
description: |
  How twitter and #tidytuesday helped me land a job offer
author:
  - name: Shannon Pileggi
date: 08-27-2020
preview: welcome_to_rstats_twitter.png
output:
  distill::distill_article:
    self_contained: false
draft: TRUE
creative_commons: CC BY
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r}
library(tidyverse)
```


# TL; DR

Weaving resources from #tidytuesday, twitter, and blog posts, I challenged myself to create this [interview presentation](https://spileggi.shinyapps.io/Pileggi_presentation_shiny/) to gain insights on my prospective employer while showcasing my analytics abilities.

# Back up

When I was told that my interview at [Adelphi Research](https://adelphiresearch.com/) would involve a presentation, I was pretty happy. A presentation allows the candidate to highlight what they choose, and, after being a college professor for six years, I was confident in my public speaking. The specific presentation instructions were:

> This can be any 20 minute presentation you’d like to share with the team; our recommendation is that it focuses on Market Research, is something innovative, and something that you’re particularly proud of!

It was also briefly mentioned in the phone screening that the analytics group with whom I was interviewing was just beginning to develop [shiny](https://shiny.rstudio.com/) applications.  

# Formulation

Then I faced the dilemma of what to actually present. I wanted to shine as a candidate, but my previous academic experience was tangential to this industry. I took their instructions to heart and decided to create something that I was truly proud of. I developed a project with three goals in mind:

 1. Learn about my prospective employer.
 
 2. Cultivate my nascent tidyverse skills.
 
 3. Invest in new analytic methods and coding techniques that could serve future me.  

# Fortuitous tweets

At the same time, 2018 was the inaugural year of [#tidytuesday](https://github.com/rfordatascience/tidytuesday). Prompted by [David Robinson](https://twitter.com/drob)'s [screencast](https://www.youtube.com/watch?v=C69QyycHsgE) analyzing medium articles with [tidytext](https://github.com/juliasilge/tidytext) I dove into my first foray in text analysis, which inspired me to apply the same principles to tweets.

```{r}
tweetrmd::include_tweet("https://twitter.com/drob/status/1069954125020958720")
```

I was also concerned with the actual format / deliverable for the presentation. I knew I wanted to create something in R, but that [xaringan](https://bookdown.org/yihui/rmarkdown/xaringan.html) might not work well for self-contained presentations that needed to be emailed around. I tweeted an #rstats plea for advice and was inspired by [Emily Riederer](https://twitter.com/EmilyRiederer)'s suggestion to use a [flexdashboard](https://rmarkdown.rstudio.com/flexdashboard/index.html) for the presentation.  


```{r}
tweetrmd::include_tweet("https://twitter.com/EmilyRiederer/status/1073001289339027456")
```


# What I did

I spent many hours at the computer. I paused and replayed David Robinson's screencast several times. I researched R functions and analytic methods that were new to me.  And then...  
 
I learned about my prospective employer through their tweets. I explored their tweets over time, common hashtags, and tweet popularity by likes and retweets.  I examined trends in hashtags over time, the distribution of likes and retweets among common hashtags, and a network analysis of what hashtags went together. I also delved into common words (but not hashtags) in the tweets, a network analysis of those words, and a lasso model to assess word associations with likes and retweets.  Lastly, I created two shiny apps to accompany the presentation, one to search tweets and another to explore the network analysis of words and hashtags. Yes, it was a lot! (And my presentation did go slightly over 20 minutes. `r emo::ji("grimace")`)


# R packages    

Starred packages (`*`) are packages that I used for the first time (7!). Unfortunately, I no longer have links to the wonderful blog posts that helped me navigate these packages.

```{r}
learned <- tibble::tribble(
  ~Package, ~Purpose,
  "*[rtweet](https://github.com/ropensci/rtweet)", "retrieve tweets",
  "*[tidytext](https://github.com/juliasilge/tidytext)", "manipulate tweet text",
  "[lubridate](https://lubridate.tidyverse.org/)", "manipulate tweet dates",
  "[dplyr](https://dplyr.tidyverse.org/)", "manipulate data", 
  "[stringr](https://stringr.tidyverse.org/)", "re-code hashtags", 
  "[skimr](https://github.com/ropensci/skimr)", "preview data", 
  "[forcats](https://forcats.tidyverse.org/)", "ordering factors", 
  "[ggplot2](https://ggplot2.tidyverse.org/)", "data visualization", 
  "*[widyr](https://github.com/dgrtwo/widyr)", "pairwise correlations", 
  "*[igraph](https://igraph.org/r/doc/)", "create network", 
  "*[ggraph](https://ggraph.data-imaginist.com/index.html)", "visualize network", 
  "*[glmnet](https://glmnet.stanford.edu/)", "lasso model",
  "[broom](https://github.com/tidymodels/broom)", "tidy model output",
  "[DT](https://rstudio.github.io/DT/)", "display tables",
  "*[flexdashboard](https://rmarkdown.rstudio.com/flexdashboard/)", "format presentation",
  "*[shiny](https://shiny.rstudio.com/)", "create shiny app"
  )
```



```{r}
learned %>% 
 AdelphiR::my_kable()
```



# Results

I hosted my presentation and the two shiny applications on [shinyapps.io](https://www.shinyapps.io/) and made all of my code publicly available on github. 

  * [Interview presentation](https://spileggi.shinyapps.io/Pileggi_presentation_shiny/) 
  
  * [Shiny app for tweet look up](https://spileggi.shinyapps.io/tweet_lookup/)  
  * [Shiny app for network analysis](https://spileggi.shinyapps.io/tweet_network/)
  
  * [Code on github](https://github.com/shannonpileggi/Adelphi-tweets) behind it all

I got the job!  Along the way, I gleaned insights about the values and industry of my prospective employer, empowering me throughout the interview process. In addition, my interviewers forwarded the presentation to their PR team to discuss their twitter presence. 

Yes, I probably could have secured a job offer with less effort, but I would have been less confident. I embraced the interview as an opportunity to create my own side project to foster new skills that would serve me for other potential interviews, personal projects, or work projects. And enabled by the #rstats twitter presence, the incredible packages that I used, and the numerous blog posts that I referenced, I had fun. Cheers, R! `r emo::ji("cheers")`   


# Acknowledgements

Thumbnail artwork by [`@allison_horst`](https://twitter.com/allison_horst?lang=en).  Thanks to [Megan McClintock](https://www.linkedin.com/in/megan-mcclintock/) (my sister) for suggestions.


