---
title: "What's in a name?"
author: "Shannon Pileggi"
date: 2018-10-16
categories: ["R"]
tags: ["blogs", "piping", "tidyverse", "hot data"]
---

So do I have **piping hot** _data_ or am I **piping** _hot data_?  Let's break it down.

**Piping Hot** _Data_

Wow, my _data_ is **piping hot**!  This connotes exciting, newly released data!  I can't promise that I'll fulfill this expectation.  Let's just say that I'll talk about data occasionally, and on special occassions it might even be piping hot.

**Piping** _Hot Data_

Here, I am **piping** my elusive _hot data_. This is what I was really going for - an ode to the **pipe** in R:  

> `%>%`

The [pipe operator](http://r4ds.had.co.nz/pipes.html) simplifies long operations by [linking multiple functions]((https://www.datacamp.com/community/tutorials/pipe-r-tutorial)) simultaneously.  Although the coding construct of the pipe has been floating around since the 1970's, the floodgates didn't open for R until [2014](http://adolfoalvarez.cl/plumbers-chains-and-famous-painters-the-history-of-the-pipe-operator-in-r/).  I don't know the exact date, but I do remember the first time I saw those three characters and the sea of emotions that rained down.  Confusion. Curiousity. Excitement. 

In just 4 short years, the pipe and its friends in the [tidyverse](https://www.tidyverse.org/) have revolutionized how we code in R, to the point that you may feel illiterate at conferences if you don't have some baseline understanding - _at first_.  Because the beauty of the pipe is that it streamlines readibility of R code, such that even if you have never done it, you can still get the gist of what is going on.  So much so that believers are proselytizing ["Teach the tidyverse to beginners"](http://varianceexplained.org/r/teach-tidyverse/)!

Let's lay the pipelines with a quick example using the classic `iris` data set.  To get started, load the `tidyverse` library and get an overivew of the data.

```{r message = F, start}
library(tidyverse)

glimpse(iris)
```

Our simple objective is to compute the mean `Sepal.Length` for each `Species` in the data set.  There are *many* ways to accomplish this without the tidyverse, but for the sake of side-by-side comparisons, I'll first demo using tidyverse functions without piping.  This is blasphemy. Don't ever do this.

```{r, nopipe}
grp_iris <- group_by(iris, Species)

mean_by_species <- summarise(grp_iris, mean = mean(Sepal.Length))

mean_by_species
```

With this method I created an intermediate object (`grp_iris`) in order to get the final desired result.  And it's not easy to follow how the objects relate: note how you define objects on the left hand side of `->` and then leapfrog them over to the right hand side of `->`, continuing this sequence in perpetuity for all of your operations.  That looks like some plumbing prone to leaks.

Let's try the same manipulations utilizing piping.

```{r, withpipe}
iris %>% glimpse

mean_by_species <- iris %>% 
                     group_by(Species) %>% 
                     summarise(mean = mean(Sepal.Length))
mean_by_species
```

Voilá!  It's clear from the left side of the pipe that all manipulations are done on the `iris` data set, and it's clear from the right side of the pipe that the series of operations performed are `group_by` and `summarise`.  Wow, I like the way data flow through those pipes!    

I'm still not as not as fluent with piping in the tidyverse as I'd like to to be. Sometimes I feel like I am chattering in a second language rather than speaking my mother tongue.  So here are my goals for **_Piping Hot Data_**:

* Pipe my way through the tidyverse.
* Demo data science tools and methods. 
* Discover new data and R packages.
* Deliberate data and technical topics.


I hope you enjoy!

